Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-05-17 paper.tex
Title change, comments, edits.

\title{Inferring Graphics Programs from Images}

@@

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a neural network that proposes plausible drawing primitives
  that explain an image.  From these deep net proposals we can use
  program synthesis techniques to recover a graphics
  program with constructs like variable bindings, iterative loops, or
  simple kinds of conditionals. With a graphics program in hand,
  we can extrapolate figures, form visual analogies, \remark{what else should go here?}, and correct errors made by the deep network.
  Taken together these results are a step towards
  agents that induce programs from perceptual input.  
\end{abstract}

@@

 How could an agent go from noisy, high-dimensional perceptual input
 to a symbolic, abstract object, like a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting natural images, such as hand
 drawings, into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Fig.\ref{firstPageExamples}).
 %% \remark{The use of `graphics programs / visual programs' in the paper
 %%   title, title of this section, and the body of this section feels
 %%   too broad. `Graphics program' could conjur a lot of different ideas
 %%   (esp. 3D graphics); don't want to set the reader up to expect one
 %%   thing and then be disappointed that what you've done isn't
 %%   that. You bring up diagram-drawing later in the intro; I think it
 %%   should be made clear sooner (and certainly mentioned explicitly in
 %%   the abstract, when you get around to writing that).}

\newcommand{\exampleImageSize}{2cm}
\begin{figure}[h]
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \texttt{for} $0\leq i < 4$:\\
    \texttt{rectangle} $(-3i + 9,2i - 2, -3i + 11,6)$\\
    \texttt{for} $0\leq j < i$:\\
    \texttt{circle} $(3i-2,-2j+5)$\\
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(bottom). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
  \end{figure}

High dimensional perceptual input may seem ill matched to the abstract
semantics of a programming language. But programs with constructs like
recursion or iteration produce a simpler \emph{execution trace} of
primitive actions; for our domain the primitive actions are drawing
commands. Our hypothesis is that the execution trace of the program is
better aligned with the perceptual input, and that the trace can act
as a kind of bridge between perception and programs. We test this
hypothesis by developing a model that learns to map from an image to
the execution trace of the graphics program that drew it.  With the
execution trace in hand, we can bring to bear techniques from the
program synthesis community to recover the latent graphics program.
This family of techniques, called \emph{constraint-based program synthesis}~\cite{solar2008program},
work by modeling a set of possible programs inside of a constraint solver,
like a SAT or SMT solver~\cite{de2008z3}.
These techniques excel at uncovering high-level symbolic structure,
but are not well equipped to deal with real-valued perceptual inputs.

We develop a hybrid architecture for inferring graphics programs.  Our
approach uses a deep neural network infer an execution trace from an
image; this network recovers primitive drawing operations such as
lines, circles, or arrows, along with their parameters. For added
robustness, we use the deep network as a proposal distribution for a
stochastic search over execution traces.  Finally, we use techniques
in the program synthesis community to recover the program from its
trace.  The program synthesizer discovers constructs like loops and
geometric operations like reflections and affine transformations.
\remark{This paragraph is all about making things a bit more specific,
  so you really need more specifics about program synth here.}

Each of these three components -- the deep network, the stochastic
search, the program synthesizer -- confers its own advantages. From
the deep network, we get a fast system that can recover plausible
execution traces in about a minute~\remark{A minute seems slow to me, for deep net inference. Are you talking about training time, here, or...?}. From the stochastic search we get
added robustness; essentially, the stochastic search can correct
mistakes made by the deep network's proposals.  From the program
synthesizer, we get abstraction: our system recovers coordinate
transformations, for loops, and subroutines, which are useful for
downstream tasks and can help correct some mistakes of the earlier stages.
\remark{I wonder if this would work even better as a bulleted list...}

@@

In the computer graphics literature, there have been other systems which convert sketches into procedural representations. One uses a convolutional network to match a sketch to the output of a parametric 3D modeling system~\cite{huang2017shape}. Another uses convolutional networks to support sketch-based instantiation of procedural primitives within an interactive architectural modeling system~\cite{Nishida:2016:ISU:2897824.2925951}. Both systems focus on inferring fixed-dimensional parameter vectors. In contrast, we seek to automatically learn a structured, programmatic representation of a sketch which captures higher-level visual patterns.

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far to a convolutional network -- these two inputs are represented as
separate channels for the convnet. Given the features extracted by the
convnet, a multilayer perceptron then predicts a distribution over the
next drawing command to add to the trace.  We predict the drawing
command token-by-token, and condition each token both on the image
features and on the previously generated tokens.  For example, the
network first decides to emit the \verb|circle| token conditioned on
the image features, then it emits the $x$ coordinate of the circle
conditioned on the image features and the \verb|circle| token, and
finally it predicts the $y$ coordinate of the circle conditioned on
the image features, the \verb|circle| token, and the $x$ coordinate.
\remark{There are some more details that are important to provide
  about this architecture, though possibly in an Appendix: the
  functional form(s) of the probability distributions over tokens, the
  network layer sizes, which MLPs share parameters, etc.}

The distribution over the next drawing command factorizes:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta [t_k | f_\theta(I,\text{render}(T)), \{t_j\}_{j = 1}^{k - 1}]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, and $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network). The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, and the \verb|STOP|
token is emitted by the network to signal that the execution trace
explains the image.

We train the network by sampling execution traces $T$ and target
images $I$ for randomly generated scenes, and maximizing
(\ref{objective}) wrt $\theta$ by gradient ascent.
Training does not require backpropagation across the entire sequence of drawing commands:
drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.
In a sense, this model is like an autoregressive variant of AIR~\cite{eslami1603attend} without attention.

@@

This network suffices to ``derender'' images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a SMC sampling scheme.  For
the SMC sampler, we use pixel wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathcal{R}$ uses the distance between two images as a proxy for a
likelihood.

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (figure~\ref{handDrawingExamples}).
Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learned} a surrogate likelihood function.
Our learned $L(\cdot|\cdot)$
is a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
Formally we train our likelihood surrogate to approximate:
\begin{equation}
-\log L(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|
  \end{equation}

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn reasonably carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid,
we made the drawings on graph paper.
\begin{figure}[H]%
  \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}%
   \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/60-2-reduced.png} 
    \subcaption{}
  \end{minipage}
  \caption{(a): a hand drawing. (b): Rendering of the parse our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) \& (d) for noisy renderings of (b).}\ref{handDrawingExamples}
\end{figure}

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-0.png}
    \includegraphics[width = 2cm]{figures/0-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}    
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}}\label{handDrawingExamples}
  \end{figure}

\section{Synthesizing graphics programs from execution traces}
Although the execution trace of a graphics program describes the parts
of a scene, it fails to encode higher-level features of the image,
such as repeated motifs, symmetries or reflections.  A \emph{graphics
  program} better describe structures like these,
and we now take as our goal to synthesize simple graphics programs from
their execution traces.

We constrain the space of allowed programs by writing down a context
free grammar over a space of programs. Although it might be desirable
to synthesize programs in a Turing-complete language like Lisp or
Python, a more tractable approach is to specify what in the program
languages community is called a Domain Specific Language (DSL). Our DSL (Table~\ref{DSL})
encodes prior knowledge of what graphics programs tend to look like.

@@

Given the DSL and a trace $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{\substack{p\in \text{DSL}\\p \text{ evaluates to } T}} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of statements it contains, where a
statement is a ``Command'' in Table~\ref{DSL}.
We also penalize using more distant constants,
which encourages reuse of coefficients.
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

@@

Why synthesize a graphics program,
if the execution trace already suffices to recover the objects in an image?
Within our domain of hand-drawn figures, graphics program synthesis has several uses:
\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we can change all of the circles to squares,
were make all of the lines be dashed.
We can also \textbf{extrapolate} figures
by increasing the number of times that loops are executed.

@@

\subsection{Correcting errors made by the neural network}
The program synthesizer can help correct errors from  the neural network by favoring execution traces which lead to more concise or general programs.
For example, one generally prefers figures with neatly aligned objects over figures whose parts are slightly misaligned -- and nice alignment lends itself to short programs.
Similarly, figures often have repeated parts,
which the program synthesizer might model as a reflection.
So, if one is considering several candidate traces proposed by the neural network,
we might prefer traces whose best programs have desirable features like being short or having iterated structures.

Concretely we implemented the following scheme: the neurally guided sampling scheme of section~\ref{neuralNetworkSection} for image $I$ produces (samples) candidate traces, $\mathcal{F}(I)$.
Instead of predicting the most likely trace in $\mathcal{F}(I)$ according to the neural network,
we can take into account the programs that best explain the traces. 
Writing $\hat{T}(I)$ for the trace the model predicts for image $I$,
\begin{equation}
\hat{T}(I) = \argmax_{T\in \mathcal{F}(I)} L(I | \text{render}(T)) \times\probability_{\beta} [ \text{program}(T)] 
\end{equation}
where $\probability_{\beta} [\cdot]$ is a prior probability
distribution over programs parameterized by $\beta$.
This is equivalent to doing
MAP inference in a generative model where the program is first drawn
from $\probability_{\beta} [\cdot]$, then the program is executed deterministically,
and then we observe a noisy version of the program's output, where $L$
is the noise model.

Given a corpus of graphics program synthesis problems with annotated ground truth traces (so pairs of $(I,T)$),
we find a maximum likelihood estimate of $\beta$:
\begin{equation}
  \beta^* = \argmax_{\beta} \expect \left[ \log \frac{\probability_{\beta} [\text{program}(T)]\times L(I|\text{render}(T))}{\sum_{T'\in \mathcal{F}(I)} \probability_{\beta} [\text{program}(T')]\times L(I|\text{render}(T'))} \right]
\end{equation}
where the expectation is taken both over the model predictions and the
$(I,T)$ pairs in the training corpus.  We define $\probability_{\beta}
[\cdot]$ to be a log linear distribution $\propto \exp
(\beta\cdot \phi(\text{program}))$, where $\phi(\cdot)$ is a feature
extractor for programs.  We can extract a few basic features of a
program, like its size or how many loops it has, and use these
features to help predict whether a trace is the correct explanation
for an image.

>>>>

% \title{Inferring Graphics Programs from Images}
\title{Learning to Infer Graphics Programs from Hand-Drawn Images}

@@

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing primitives
  that explain an image.  From these deep net proposals we can use
  program synthesis techniques to recover a graphics
  program with constructs like variable bindings, iterative loops, or
  simple kinds of conditionals.~\remark{Can you (concisely) communicate the idea of treating the deep net proposal as an `execution trace?' It's a really neat intellectual framing that deserves to be foregrounded.} With a graphics program in hand,
  we can correct errors made by the deep network, extrapolate drawings, and~\remark{what else: `visual analogies' is vague. `Cluster drawings by use of similar high-level program structures?'}
  Taken together these results are a step towards
  agents that induce useful, human-readable programs from perceptual input.  
\end{abstract}

@@

 How can an agent convert noisy, high-dimensional perceptual input
 to a symbolic, abstract object, such as a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting natural images, such as hand
 drawings, into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Figure~\ref{firstPageExamples}).
 %% \remark{The use of `graphics programs / visual programs' in the paper
 %%   title, title of this section, and the body of this section feels
 %%   too broad. `Graphics program' could conjur a lot of different ideas
 %%   (esp. 3D graphics); don't want to set the reader up to expect one
 %%   thing and then be disappointed that what you've done isn't
 %%   that. You bring up diagram-drawing later in the intro; I think it
 %%   should be made clear sooner (and certainly mentioned explicitly in
 %%   the abstract, when you get around to writing that).}

\newcommand{\exampleImageSize}{2cm}
\begin{figure}[h]
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \texttt{for} $0\leq i < 4$:\\
    \texttt{rectangle} $(-3i + 9,2i - 2, -3i + 11,6)$\\
    \texttt{for} $0\leq j < i$:\\
    \texttt{circle} $(3i-2,-2j+5)$\\
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(bottom). (b) Synthesizes high-level \emph{graphics program} from hand drawing.~\remark{Any objection to putting this on page bottom?}}\label{firstPageExamples}
  \end{figure}

High dimensional perceptual input may seem ill matched to the abstract
semantics of a programming language. But programs with constructs such as
recursion or iteration produce a simpler \emph{execution trace} of
primitive actions; for our domain, the primitive actions are drawing
commands. Our hypothesis is that the execution trace of the program is
better aligned with the perceptual input, and that the trace can act
as a kind of bridge between perception and programs. We test this
hypothesis by developing a model that learns to map from an image to
the execution trace of the graphics program that drew it.  With the
execution trace in hand, we can bring to bear techniques from the
program synthesis community to recover the latent graphics program.
This family of techniques, called \emph{constraint-based program synthesis}~\cite{solar2008program},
work by modeling a set of possible programs inside of a constraint solver,
such as a SAT or SMT solver~\cite{de2008z3}.
These techniques excel at uncovering high-level symbolic structure,
but are not well equipped to deal with real-valued perceptual inputs.

We develop a hybrid architecture for inferring graphics programs.  Our
approach uses a deep neural network infer an execution trace from an
image; this network recovers primitive drawing operations such as
lines, circles, or arrows, along with their parameters. For added
robustness, we use the deep network as a proposal distribution for a
stochastic search over execution traces.  Finally, we use program synthesis to recover the program from its
trace.  The program synthesizer discovers constructs such as loops and
geometric operations such as reflections and affine transformations.
% \remark{This paragraph is all about making things a bit more specific,
%   so you really need more specifics about program synth here.}

Each of these three components -- the deep network, the stochastic
search, the program synthesizer -- confers its own advantages. From
the deep network, we get a fast system that can recover plausible
execution traces in about a minute~\remark{A minute seems slow to me, for deep net inference...}. From the stochastic search, we get
added robustness: essentially, the stochastic search can correct
mistakes made by the deep network's proposals.  From the program
synthesizer, we get abstraction: our system recovers coordinate
transformations, for loops, and subroutines, which are useful for
downstream tasks and can help correct some mistakes of the earlier stages.
% \remark{I wonder if this would work even better as a bulleted list...}
\remark{Tie these claims into the paper results: state what the `downstream' tasks are that you actually do, and refer to the sections where they occur}

@@

In the computer graphics literature, there have been other systems which convert sketches into procedural representations. One uses a convolutional network to match a sketch to the output of a parametric 3D modeling system~\cite{huang2017shape}. Another uses convolutional networks to support sketch-based instantiation of procedural primitives within an interactive architectural modeling system~\cite{Nishida:2016:ISU:2897824.2925951}. Both systems focus on inferring fixed-dimensional parameter vectors. In contrast, we seek to automatically infer a structured, programmatic representation of a sketch which captures higher-level visual patterns.

@@

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given the features extracted by the
convnet, a multilayer perceptron then predicts a distribution over the
next drawing command to add to the trace.  We predict the drawing
command token-by-token, conditioning each token both on the image
features and on the previously generated tokens.  For example, the
network first decides to emit the \verb|circle| token conditioned on
the image features, then it emits the $x$ coordinate of the circle
conditioned on the image features and the \verb|circle| token, and
finally it predicts the $y$ coordinate of the circle conditioned on
the image features, the \verb|circle| token, and the $x$ coordinate.
\remark{There are some more details that are important to provide
  about this architecture in the supplement: the
  functional form(s) of the probability distributions over tokens, the
  network layer sizes, which MLPs share parameters, etc.}

The distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta [t_k | f_\theta(I,\text{render}(T)), \{t_j\}_{j = 1}^{k - 1}]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, and $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network). The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, and the \verb|STOP|
token is emitted by the network to signal that the execution trace
explains the image.~\remark{Make explicit that a $T_n$ in Equation 2 is a concise way of referring to a sequence of tokens from Equation 1?}

We train the network by sampling execution traces $T$ and target
images $I$ for randomly generated scenes~\remark{this process ought to be explained, perhaps in supplement if it is at all detailed}, and maximizing
(\ref{objective}) with respect to $\theta$ by gradient ascent.
Training does not require backpropagation across the entire sequence of drawing commands:
drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.
In a sense, this model is like an autoregressive variant of AIR~\cite{eslami1603attend} without attention.
~\remark{Somewhere (not necessarily here) you should probably cite the deep learning toolkit you used.}

@@

This network suffices to ``derender'' synthetic images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathcal{R}$ uses the distance between two images as a proxy for a
likelihood.

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Figure~\ref{handDrawingExamples}).~\remark{Process should be described (in supplement)}
Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function.
Our learned $L(\cdot|\cdot)$
is a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings~\remark{The two $L$'s should have some kind of identifier e.g. a subscript to distinguish them}.
Formally we train our likelihood surrogate to approximate:
\begin{equation}
-\log L(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|
  \end{equation}

\remark{What distance metric is this? Some kind of edit distance?}
We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn reasonably carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid,
we made the drawings on graph paper.
\begin{figure}[H]%
  \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}%
   \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.2\textwidth}\includegraphics[width = 2cm]{figures/60-2-reduced.png} 
    \subcaption{}
  \end{minipage}
  \caption{(a): a hand drawing. (b): Rendering of the parse our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) \& (d) for noisy renderings of (b).}\ref{handDrawingExamples}~\remark{This figure is never referenced? Is the ref here supposed to be a label?}
\end{figure}

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-0.png}
    \includegraphics[width = 2cm]{figures/0-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}    
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Figure~\ref{firstPageExamples}}\label{handDrawingExamples}\remark{Showing `failure cases' (the last three above) out-of-context seems not great. Perhaps remind people in the caption that stochastic search can help correct some of these issues?}
  \end{figure}

\section{Synthesizing graphics programs from execution traces}
Although the execution trace of a graphics program describes the parts
of a scene, it fails to encode higher-level features of the image,
such as repeated motifs or symmetries.  A \emph{graphics
  program} better describe structures like these,
and we now take as our goal to synthesize simple graphics programs from
their execution traces.

We constrain the space of allowed programs by writing down a context
free grammar over a space of programs. Although it might be desirable
to synthesize programs in a Turing-complete language such as Lisp or
Python, a more tractable approach is to specify what in the program
languages community is called a Domain Specific Language (DSL)~\remark{citation?}. Our DSL (Table~\ref{DSL})
encodes prior knowledge of what graphics programs tend to look like.

@@

Given the DSL and a trace $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{\substack{p\in \text{DSL}\\p \text{ evaluates to } T}} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of statements it contains, where a
statement is a ``Command'' in Table~\ref{DSL}.
We also penalize using more distant constants~\remark{awkward},
which encourages reuse of coefficients.~\remark{More detail on this in supplement}
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

@@

Why synthesize a graphics program,
if the execution trace already suffices to recover the objects in an image?
Within our domain of hand-drawn figures, graphics program synthesis has several uses:
\remark{I'm of two minds about how these subsections should be ordered. The current ordering leads with the coolest results, which is nice. But it's a bit...deflating?...to start with such cool results and then end on the somewhat technical point of how the synthesizer can help correct parse errors. An alternative would be to flip the ordering, starting with the technical point and building toward progressively cooler results. This requires a bit more patience on the part of the reader, but the overall narrative flow/payoff feels better. In any case, whatever ordering you pick should be the same order as these applications are mentioned in the abstract.}

\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we can change all of the circles to squares or make all of the lines be dashed.
We can also \textbf{extrapolate} figures
by increasing the number of times that loops are executed.~\remark{Pick a few of these to show off and put the rest in supplement}

@@

\subsection{Correcting errors made by the parsing network}
The program synthesizer can help correct errors from the parse proposal network by favoring execution traces which lead to more concise or general programs.
For example, one generally prefers figures with perfectly aligned objects over figures whose parts are slightly misaligned -- and precise alignment lends itself to short programs.
Similarly, figures often have repeated parts,
which the program synthesizer might be able to model as a loop or reflectional symmetry.
So, in considering several candidate traces proposed by the neural network,
we might prefer traces whose best programs have desirable features such being short or having iterated structures.

Concretely, we implemented the following scheme: the neurally guided sampling scheme of section~\ref{neuralNetworkSection} for image $I$ samples candidate traces $\mathcal{F}(I)$.
Instead of predicting the most likely trace in $\mathcal{F}(I)$ according to the neural network,
we can take into account the programs that best explain the traces. 
Writing $\hat{T}(I)$ for the trace the model predicts for image $I$,
\begin{equation}
\hat{T}(I) = \argmax_{T\in \mathcal{F}(I)} L(I | \text{render}(T)) \times\probability_{\beta} [ \text{program}(T)] 
\end{equation}
where $\probability_{\beta} [\cdot]$ is a prior probability
distribution over programs parameterized by $\beta$.
This is equivalent to doing
MAP inference in a generative model where the program is first drawn
from $\probability_{\beta} [\cdot]$, then the program is executed deterministically,
and then we observe a noisy version of the program's output, where $L$
is the noise model.

Given a corpus of graphics program synthesis problems with annotated ground truth traces (i.e. $(I,T)$ pairs),
we find a maximum likelihood estimate of $\beta$:
\begin{equation}
  \beta^* = \argmax_{\beta} \expect \left[ \log \frac{\probability_{\beta} [\text{program}(T)]\times L(I|\text{render}(T))}{\sum_{T'\in \mathcal{F}(I)} \probability_{\beta} [\text{program}(T')]\times L(I|\text{render}(T'))} \right]
\end{equation}
where the expectation is taken both over the model predictions and the
$(I,T)$ pairs in the training corpus.  We define $\probability_{\beta}
[\cdot]$ to be a log linear distribution $\propto \exp
(\beta\cdot \phi(\text{program}))$, where $\phi(\cdot)$ is a feature
extractor for programs.  We extract a few basic features of a
program, such as its size and how many loops it has, and use these
features to help predict whether a trace is the correct explanation
for an image.
